### 2015-03-30 README.Rmd
small edits to readme. and moved where some files are being sourced(). model does not run: problem is that in the function PopulateNet() on line 34 of file populate_net.R a variable called 'burn.in' is passed into a function but this variable is not defined as an arg to PopulateNet() or inside PopulateNet().

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

This is code for an ABM of a climate prediction market
in which traders adapt their belief about the climate based
on the monetary performance of their neighboring traders in
a social network. The ABM is meant to test whether climate
prediction markets foster the convergence of people’s beliefs
about the climate. The model will be used to
assess whether such convergence is impacted by factors such
as the initial belief-homophily in the social network or the
time horizon at which climate securities are realized.

[Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University) and [John Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University).

Master is the main script that sources everything else.
```{r, eval=FALSE, include=TRUE}
source("master.R")
````