### 2015-08-28 Temp Latex work/Supplemental material/Supplemental material.tex
Knitr looks neat but I like my old latex editor (Texstudio) better than working from inside Rstudio to edit latex text.
Over the years, I have personalized it with shortcuts and stuff which make me more productive, and going back really slows me down.

So I am creating this directory to store temporary latex files. Once I am done with working on asection of the paper, I'll put the section back on the .Rwd file.

If there is any better way to proceed, just let me know.

\documentclass[10pt,a4paper]{article}
\usepackage{fullpage}
\usepackage[latin1]{inputenc}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{animate}
\usepackage[linktocpage, breaklinks, colorlinks = true, linkcolor = Brown, citecolor = blue]{hyperref}
\usepackage{enumerate}
\usepackage[dvips]{graphicx}
\usepackage[usenames,dvipsnames]{pstricks}
\usepackage{epsfig}
\usepackage{pst-grad} % For gradients
\usepackage{pst-plot} % For axes
\usepackage{float}
\usepackage{bm}
\usepackage{nameref}
\usepackage{xfrac}
\usepackage{accents}
\usepackage[stable]{footmisc}
\usepackage[nice]{nicefrac}
\usepackage{thmtools}
\usepackage{blindtext}
\usepackage{tcolorbox} 
\tcbuselibrary{breakable}

\newcounter{test}

\newtcolorbox{framet}[2][%
breakable,
arc=7pt,
outer arc=0pt,
coltitle=black,
fonttitle=\bfseries,
boxrule=2pt,
colframe=Brown,
colback=Apricot,
title after break={Frame~\thetest\ (Continued)}
]{%
before upper={
  \stepcounter{test}\textbf{Frame~\thetest.\ }%
},
label={#2},
#1}

\newcommand{\aligncom}[1]{%
  \text{\phantom{(#1)}} \tag{#1}
}

\newcounter{frame} \setcounter{frame}{0}

\renewcommand{\theframe}{\Roman{frame}}

\newenvironment{FFrame}[1]%
{
\refstepcounter{frame}
\vspace{1ex} \flushleft\makebox[13ex][l]{\textbf{Frame} \theframe} \texttt{#1}
}

\graphicspath{{graphics/}}

\declaretheorem [name = Theorem]{theo}
\declaretheorem [name = Lemma]{lem}
\declaretheorem [name = Axiom]{ax}
\declaretheorem [name = Preference axiom]{pref}
\declaretheorem [name = Observation]{obs}
\declaretheorem [name = Definition]{defi}
\declaretheorem [name = Example]{expl}
\declaretheorem [name = Social ordering function]{sof}
\declaretheorem [name = Problem]{prob}
\declaretheorem [name = Assumption]{ass}
\declaretheorem [name = Proposition]{prop}

\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}%
              {<-> s * [0.900] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}%
                                 {m}{it}

%\newtheorem{ax}{Axiom}
%\newtheorem{pref}{Preference axiom}
%\newtheorem{obs}{Observation}
%\newtheorem{defi}{Definition}
%\newtheorem{sof}{Social ordering function}

\newcommand{\nref}{\nameref}

% %Short commands for math mode

%Style
\newcommand{\mb}[1]{\mathbf{#1}}     % bold letters in mathmode
\newcommand{\bs}[1]{\boldsymbol{#1}} % bold symbols in mathmode
\newcommand{\bb}[1]{\mathbb{#1}}     % Doppelstrichbuchstaben im Mathematikmodus
\newcommand{\mc}[1]{\mathcal{#1}}    % calligraphy style
%Sets
\newcommand{\rr}{\bb{R}} % set of reals in mathmode
%Fractions
\newcommand{\pfraca}[2]{\bigl(\frac{#1}{#2}\bigr)}    % ratios with small parantheses
\newcommand{\pfracb}[2]{\Bigl(\frac{#1}{#2}\Bigr)}    % ratios with big parantheses
\newcommand{\pfracc}[2]{\biggl(\frac{#1}{#2}\biggr)}  % ratios with bigger parantheses
\newcommand{\pfracd}[2]{\Biggl(\frac{#1}{#2}\Biggr)}  % ratios with huge parantheses
%Partial derivatives
\newcommand{\di}[1]{\frac{\partial}{\partial #1}}     % shortcut for first partial derivative
\newcommand{\dii}[1]{\frac{\partial^2}{\partial #1^2}}% shortcut for second partial derivative
%other
\newcommand{\ra}{\rightarrow}            % convergence
%Convergence
\newcommand{\cd}{\overset{d}{\rightarrow}}    % convergence in distribution
\newcommand{\cp}{\overset{p}{\rightarrow}}    % convergence in probability
\newcommand{\cas}{\overset{as}{\rightarrow}}  % convergence almost surely
%Text in math environment
\newcommand{\sgn}{~\text{sgn}~}
\newcommand{\f}{~\text{if}~}
\newcommand{\frall}{~\text{for all}~}
\newcommand{\ow}{~\text{otherwise}~}
\newcommand{\nd}{~\text{and}~}
\newcommand{\thn}{~\text{then}~}
\newcommand{\id}{\coloneqq}
%Variant of variables (tilde, lower bar,...)
\newcommand{\lbar}[1]{\underline{#1\mkern-4mu}\mkern4mu }
\newcommand{\x}{\bm{x}}
\newcommand{\m}{\bm{m}}
\newcommand{\z}{\bm{z}}
\newcommand{\g}{\bm{\gamma}}
\newcommand{\h}{\bm{h}}
\newcommand{\y}{\bm{y}}
\newcommand{\e}{\bm{e}}
\newcommand{\la}{\bm{\lambda}}
\newcommand{\ce}{\bm{c}}
\renewcommand{\b}{\bm{\beta}}
\newcommand{\hb}{\hat{\b}}

\title{}
\author{Martin Van der Linden}

\begin{document}

\maketitle

\bibliography{library.bib}
\bibliographystyle{apalike}

\end{document}