### 2015-10-21 timeseries_example.Rmd
Updated timeseries package and climate model.

---
title: "Timeseries Example"
author: "Jonathan Gilligan"
date: "October 21, 2015"
output: html_document
html_document:
  mathjax: default
---
# Example of running the timeseries functions
\(\require{mhchem}\def\COO{\ce{CO2}}\)

## Initializing the model

The first thing we need to do is get the data and initialize models for which the true model (the physically correct one) is TSI and $\log\COO$:
```{r initialize, echo=T, eval=T}
source('prepare_data.R')
source('climate_model.R')

library(ggplot2)
library(dplyr)
library(tidyr)

theme_set(theme_bw(base_size=20))

data <- prepare_climate_data('rcp 4.5')