### 2015-10-22 timeseries/timeseries_example.Rmd
Updated interval_prob to give an option for closed or open intervals, and include NA arguments for one element of t.range, to make one-sided tests (rather than inside range).

---
html_document:
  mathjax: default
author: "Jonathan Gilligan"
date: "October 21, 2015"
output: html_document
title: "Timeseries Example"
---
# Example of running the timeseries functions
\(\require{mhchem}\def\COO{\ce{CO2}}\)

## Initializing the model

The first thing we need to do is get the data and initialize models for which the true model (the physically correct one) is Slow TSI and $\log\COO$:
```{r setup, echo=T, eval=T, message = F}
library(ggplot2)
library(dplyr)
library(tidyr)

theme_set(theme_bw(base_size=20))

source('prepare_data.R')
source('climate_model.R')

prt <- function(x,digits,flag='',mark='') {
  formatC(x, digits=digits, format='f', flag=flag, big.mark=mark)
}
```
```{r load_data, echo=T, eval=T, cache=T}
data <- prepare_climate_data('rcp 4.5')

climate_data = data$data
future_data = data$future

mdl <- new('climate_model', climate = climate_data)

max_future <- 50
history_start <- 100
```

Next, we initialize two models, one for which the true model is $\log\COO$ and the other for which the true model is Slow TSI.
We initialize each model with the first `r history_start` years of data, and in each model, we predict a true future for the next `max_future` years. This true future is based on the best linear fit to the relevant covariate for each model (Slow TSI or $\log\COO$).

```{r init_models, echo=T, eval=T, cache=T, dependson=c('load_data')}
mdl.co2.true <- init_model(mdl, n_history = history_start, n_future = max_future, true_covars = list('log.co2'), future_covars = future_data)
mdl.tsi.true <- init_model(mdl, n_history = history_start, n_future = max_future, true_covars = list('slow.tsi'), future_covars = future_data)
```

Next, we run predictions for two types of traders with each model:
```{r trader_predict, echo=T, eval=T, cache=T, dependson=c('init_models'), cache_extra = c(mdl.co2.true, mdl.tsi.true)}
today <- history_start
trader_horizon <- 10

trader.co2.true.co2 <- update_model(mdl.co2.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('log.co2'))
trader.tsi.true.co2 <- update_model(mdl.co2.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('slow.tsi'))

trader.co2.true.tsi <- update_model(mdl.tsi.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('log.co2'))
trader.tsi.true.tsi <- update_model(mdl.tsi.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('slow.tsi'))
```

Now w report the predictions:
```{r predictions, eval=T, echo=T, cache=T, dependson=c('trader_predict')}
low = 0.5
high = 0.6

predictions <- data.frame(true.model = c('log CO2', 'log CO2', 'Slow TSI', 'Slow TSI'),
                          trader.model = rep_len(c('log CO2', 'Slow TSI'), 4),
                          prob = c(
                            interval_prob(trader.co2.true.co2, n_horizon = trader_horizon, t.range = c(low, high), TRUE),
                            interval_prob(trader.tsi.true.co2, n_horizon = trader_horizon, t.range = c(low, high), TRUE),
                            interval_prob(trader.co2.true.tsi, n_horizon = trader_horizon, t.range = c(low, high), TRUE),
                            interval_prob(trader.tsi.true.tsi, n_horizon = trader_horizon, t.range = c(low, high), TRUE)
                          )
)

```

| Trader's Model | True model = $\log\COO$ | True model = Slow TSI |
|:---------------|------------------------:|------------------:|
| $\log\COO$     | `r prt(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'log CO2'],3)` | `r prt(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'log CO2'],3)` |
| Slow TSI     | `r prt(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'Slow TSI'],3)` | `r prt(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'Slow TSI'],3)` |