### 2015-12-06 README.Rmd

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

This is code for an agent-based model (ABM) of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network. The ABM is meant to test whether climate prediction markets foster the convergence of people’s beliefs about the climate.

[John J. Nay](http://johnjnay.com/) (School of Engineering, Vanderbilt University), [Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Economics, Vanderbilt University), and [Jonathan M. Gilligan](https://my.vanderbilt.edu/jonathangilligan/) (Earth and Environmental Sciences, Vanderbilt University).